---
title: "Aquatics"
engine: knitr
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(ggiraph)
library(patchwork)
library(tidyverse)
library(neon4cast)
library(score4cast)
source("R/plot-utils.R")
score4cast::ignore_sigpipe()

```

```{r include=FALSE}
cutoff <- as.character(Sys.Date() - 30)
combined <- arrow::open_dataset("cache/parquet/aquatics") |>
  filter(reference_datetime >= cutoff) |> collect()

```

## Most recent forecasts





```{r}
sites <- combined |> distinct(site_id) |> slice_head(n=6) |> collect() |> pull(site_id)
## with at least n observations to compare!
n_data <- 15
who <- combined |> filter(!is.na(observation)) |> summarise(has_data = max(reference_datetime)) |> collect()
ref <- as.character ( as.Date(who$has_data[[1]]) - n_data )
ex <- combined |> filter(reference_datetime == ref, site_id %in% sites) 
```

Below is the forecasts submitted on the date indicated below. Updates daily, selected to show forecasts for which at least some observational data has since been collected.  Mouse over to see the team id, scroll to zoom.

```{r}
ref
```



::: panel-tabset
## Oxygen

```{r}
ex |> filter(reference_datetime == ref, variable == "oxygen") |> forecast_plots()
```

## Temperature

```{r}
ex |> filter(reference_datetime == ref, variable == "temperature") |> forecast_plots()
```

## Chlorophyll-A

```{r}
ex |> filter(reference_datetime == ref, variable == "chla") |> forecast_plots()
```
:::

## Leaderboard

Average skill scores of each model across all sites. Scores are shown by reference date and forecast horizon (in days). Scores are averaged across all submissions of the model with a given horizon or a given `reference_datetime` out of submissions made since the cutoff date, `r cutoff`.


::: panel-tabset
## Oxygen

```{r}
leaderboard_plots(combined, "oxygen")
```

## Temperature

```{r}
leaderboard_plots(combined, "temperature")
```

## Chlorophyll-A

```{r}
leaderboard_plots(combined, "chla")
```
:::


## Submission statistics


```{r}
n_models <- combined |> distinct(model_id) |> nrow()
```


Between `r cutoff` and  `r Sys.Date()`:

- `r n_models` submitted to this challenge


```{r}
combined |> 
  distinct(model_id, reference_datetime) |>
  count(model_id, sort=TRUE) |>
  mutate(model_id = fct_reorder(model_id, n)) |>
  ggplot(aes(model_id, n, fill=model_id)) +
  geom_col(show.legend = FALSE) + 
  coord_flip() + 
  theme_bw() + 
  ggtitle(glue("Number of forecast submissions since {cutoff}"))
```


## Annual trends





