---
title: "Aquatics"
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(ggiraph)
library(patchwork)
library(tidyverse)
library(neon4cast)
library(score4cast)
source("R/plot-utils.R")
score4cast::ignore_sigpipe()

```

```{r include=FALSE}
cutoff <- as.character(Sys.Date() - 30)
combined <- arrow::open_dataset("cache/parquet/aquatics") |>
  filter(date >= cutoff) |> collect()

```

## Most recent forecasts

Below is the latest forecasts for which we have at least 10 observations. Mouse over to see the team id, scroll to zoom.

```{r}
sites <- combined |> distinct(site_id) |> slice_head(n=6) |> collect() |> pull(site_id)
## with at least n observations to compare!
n_data <- 10
who <- combined |> filter(!is.na(observation)) |> summarise(has_data = max(reference_datetime)) |> collect()
ref <- as.character ( as.Date(who$has_data[[1]]) - n_data )
ex <- combined |> filter(reference_datetime == ref, site_id %in% sites) 

```

::: panel-tabset 

## Oxygen

```{r}
ex |> filter(reference_datetime == ref, variable == "oxygen") |> forecast_plots()
```

## Temperature

```{r}
ex |> filter(reference_datetime == ref, variable == "temperature") |> forecast_plots()
```

## Chlorophyll-A


```{r}
ex |> filter(reference_datetime == ref, variable == "chla") |> forecast_plots()
```


:::



## Leaderboard


Average skill scores of each model across all sites.  
Scores are shown by reference date and forecast horizon (in days).  
Scores are averaged across all submissions of the model with a given horizon or a given `reference_datetime` out of submissions made since the cutoff date:

```{r include=TRUE}
cutoff
```

::: panel-tabset

## Oxygen

```{r}
leaderboard_plots(combined, "oxygen")
```

## Temperature

```{r}
leaderboard_plots(combined, "temperature")
```

## Chlorophyll-A

```{r}
leaderboard_plots(combined, "chla")
```
:::
