---
title: "Aquatics"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r setup}
library(ggiraph)
library(patchwork)
library(tidyverse)
library(neon4cast)
library(score4cast)


library(thematic)
thematic_rmd(bg="white", fg="black", accent="blue")

source("R/plot-utils.R")


```

```{r include=FALSE}
dashboard <- yaml::read_yaml("dashboard.yml") # Some configurable options

combined <- neon4cast::combined_scores("aquatics") |>
  filter(time >= lubridate::as_datetime(dashboard$start_time)) 

```

## Most recent forecasts

Below is the latest forecasts for which we have at least 15 observations. Mouse over to see the team id, scroll to zoom.

```{r}
## date of each team's most recent forecast
days <- combined |> group_by(model_id) |> summarise(date = max(start_time))

## date of each team's most recent forecast for which data is available
combined |> filter(!is.na(observed)) |>
  count(target_id, variable, model_id, start_time, site_id) |>
  filter(n > 15) |> ungroup() |>
  group_by(model_id) |> summarise(date = max(start_time))


## last date for which data is available (end of the previous month)
combined |>  filter(!is.na(observed)) |>  filter(time == max(time)) |>  summarise(date = max(time))


day <- combined |>
  filter(!is.na(observed)) |>
  count(target_id, variable, model_id, start_time, site_id) |>
  filter(n > 15) |> 
  # at least 15 data points
  filter(start_time == max(start_time, na.rm = TRUE)) |> # must happen second
  select(start_time) |> distinct(start_time) |> first()
```

::: panel-tabset
## Oxygen

```{r}
combined |> filter(start_time == day, variable == "oxygen") |> forecast_plots()
```

## Temperature

```{r}
combined |> filter(start_time == day, variable == "temperature") |> forecast_plots()
```

## Chlorophyll-A

```{r}
# combined |> filter(start_time == day, variable == "chla") |> forecast_plots()
```
:::

```{r}
# no persistence submissions...
filled_scores <- score4cast::fill_scores(combined, "climatology")
```

```{r}
leaderboard <-  filled_scores |> 
  group_by(variable, model_id) |>
  summarise(crps = mean(crps, na.rm=T),
            logs = mean(logs, na.rm=T),
            percent_na = mean(is.na(crps_model)),
            .groups = "drop") 
```

<!-- Heading 1: navbar -->

::: panel-tabset
## Oxygen

```{r}
leaderboard_plots(leaderboard, NULL, "oxygen")
```

## Temperature

```{r}
leaderboard_plots(leaderboard, NULL, "temperature")
```

## Chlorophyll-A

```{r}
leaderboard_plots(leaderboard, NULL, "chla")
```
:::

------------------------------------------------------------------------

Overall team rankings by forecast skill, as well as skill by date of initial forecast.

**Top plots**: Forecast skill measured by CRPS and log skill. Lower values indicate better predictions. Both score probabilistic forecasts, but log skill penalizes observations outside expected range much more heavily.

**Lower plot**: Forecast skill over time at each site. Not all teams submit multiple forecasts.

**Tip**: *Mouse over a color to highlight the scores of one team*
