---
title: "Beetles"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

library(ggiraph)
library(patchwork)
library(tidyverse)
library(neon4cast)
library(score4cast)
```

```{r include=FALSE, cache=TRUE}
combined <- combined_scores("beetles")
filled_scores <- fill_scores(combined, "EFI_avg_null")
```

```{r}

## problems with matching ISOweek dates mean no scores...

leaderboard <-  filled_scores %>% 
  group_by(variable, model_id) %>%
  summarise(crps = mean(crps, na.rm=TRUE),
            logs = mean(logs, na.rm=TRUE),
            percent_na = mean(is.na(crps_model))) %>% 
  arrange(crps)
#leaderboard %>% DT::datatable(fillContainer = FALSE, escape=FALSE)
```

<!-- Heading 1: navbar -->

```{r}
board1 <- 
leaderboard %>% filter(variable == "richness") %>%
  pivot_longer(cols = c(crps, logs), names_to="metric", values_to="score") %>%
  ggplot(aes(model_id, score, fill=model_id, col=model_id)) + 
  geom_col_interactive(aes(tooltip = model_id, data_id = model_id), show.legend = FALSE) + 
  facet_wrap(~metric, scales = "free_y") + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) + 
  ggtitle("richness forecasts")
```

```{r}

by_start <-  filled_scores %>% 
  filter(variable == "richness")  %>% 
  group_by(variable, model_id, start_time) %>%
  summarise(crps = mean(crps),
            logs = mean(logs),
            percent_na = mean(is.na(crps_model))) %>% 
  arrange(crps)

board2 <- 
  by_start %>% 
  ggplot(aes(start_time, crps, col=model_id)) + 
    geom_point_interactive(aes(tooltip = model_id, data_id = model_id), size=2, show.legend = FALSE) +
    geom_line_interactive(aes(tooltip = model_id, data_id = model_id), size=1, show.legend = FALSE)
```

## Row {data-height="650"}

### Species Richness: forecast skill

```{r girafe, fig.width=8, fig.height=4}
ggob <- board1 / board2 # patchwork stack
girafe(ggobj = ggob,
       width_svg = 8, height_svg = 4,
  options = list(
    opts_hover_inv(css = "opacity:0.20;"),
    opts_hover(css = "stroke-width:2;"),
    opts_zoom(max = 4)
  ))
```

------------------------------------------------------------------------

Overall model_id rankings by forecast skill, as well as skill by date of initial forecast.

**Top plots**: Forecast skill measured by CRPS and log skill. Lower values indicate better predictions. Both score probabilistic forecasts, but log skill penalizes observations outside expected range much more heavily.

**Lower plot**: Forecast skill over time at each site. Not all teams submit multiple forecasts.

**Tip**: *Mouse over a color to highlight the scores of one a team*
