---
title: Phenology 
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r setup}
library(ggiraph)
library(patchwork)
library(tidyverse)
library(neon4cast)
library(score4cast)
library(thematic)
thematic_rmd(bg="white", fg="black", accent="blue")

```

```{r}
# load the relevant data, and compute filled_scores

dashboard <- yaml::read_yaml("dashboard.yml") # Some configurable options

#combined <- combined_scores("phenology")
combined <- arrow::open_dataset("cache/parquet/phenology") |>
  filter(datetime >= lubridate::as_datetime(dashboard$reference_datetime)) |>
  collect() |> score4cast::include_horizon()
```

## Most recent forecasts

Below is the latest forecasts for which we have at least 15 observations. Mouse over to see the team id, scroll to zoom.

```{r}
green <- combined |>
  filter(variable == "gcc_90") 
sites <- green |> select(site_id) |> distinct() |> pull(site_id)
day <- green |>
  filter(!is.na(observation), site_id == sites[[1]]) |>
  count(variable, model_id, reference_datetime) |>
  filter(n > 15) |> # at least 10 data points
  filter(reference_datetime == max(reference_datetime)) |> # must happen second
  select(reference_datetime) |> distinct(reference_datetime) |> first()
```

```{r}

plotme <- function(df) {
  ggobj <- 
    ggplot(df) +
    geom_point(aes(datetime, observation)) + 
    geom_ribbon_interactive(aes(x = datetime, ymin = quantile02.5, ymax = quantile97.5,
                                fill = model_id, data_id = model_id, tooltip = model_id),
                            alpha = 0.2, show.legend=FALSE) +
    geom_line_interactive(aes(datetime, mean, col = model_id, 
                              tooltip = model_id, data_id = model_id), show.legend=FALSE) +
    facet_wrap(~site_id, scales = "free_y") +
    labs(y = "greenness index (gcc 90)") +
    theme(axis.text.x = element_text( angle = 90, hjust = 0.5, vjust = 0.5)) + 
    theme_bw()
    
  
  girafe(ggobj = ggobj,
       width_svg = 8, height_svg = 4,
  options = list(
    opts_hover_inv(css = "opacity:0.20;"),
    opts_hover(css = "stroke-width:2;"),
    opts_zoom(max = 4)
  ))

}
```

::: panel-tabset
## Sites 1 - 6

```{r}
green |> filter(reference_datetime == day) |> filter(site_id %in% sites[1:6]) |> plotme()
```

## 7-12

```{r}
green |> filter(reference_datetime == day, site_id %in% sites[7:12]) |> plotme()
```

## 13 - 18

```{r}
green |>  filter(reference_datetime == day, site_id %in% sites[13:18]) |>  plotme()
```
:::

## Leaderboard

```{r}
filled_scores <- fill_scores(combined, null="persistence")
```

```{r}
## summary statistics
leaderboard <-  filled_scores |> 
  group_by(variable, model_id) |>
  summarise(crps = mean(crps),
            logs = mean(logs),
            percent_na = mean(is.na(crps_model)),
            .groups = "drop") |> 
  mutate(model_id = fct_rev(fct_reorder(model_id, crps)))  # sort by score
```

<!-- Heading 1: navbar -->

```{r fig.height=10}
board1 <- 
leaderboard |>
  pivot_longer(cols = c(crps, logs), names_to="metric", values_to="score") |>
  ggplot(aes(x = model_id, y= score, fill=model_id, col=model_id)) + 
  geom_col_interactive(aes(tooltip = model_id, data_id = model_id),
                       show.legend = FALSE) + 
  coord_flip() +
  facet_wrap(~metric, scales='free') + 
  theme(axis.text.x = element_blank()) +
  theme_bw()
```

```{r}

by_start <-  filled_scores |> 
  group_by(variable, model_id, reference_datetime) |>
  summarise(crps = mean(crps),
            logs = mean(logs),
            percent_na = mean(is.na(crps_model))) |> 
  arrange(crps)

board2 <- 
  by_start |> 
  filter(percent_na < .1) |>
  ggplot(aes(reference_datetime, crps, col=model_id)) + 
  geom_line_interactive(aes(tooltip = model_id, data_id = model_id), size=1, show.legend = FALSE)  +
  labs(title = "Forecast skill") +
  theme_bw()

board2_logs <- 
  by_start |> 
  filter(percent_na < .1) |>
  ggplot(aes(reference_datetime, logs, col=model_id)) + 
  geom_line_interactive(aes(tooltip = model_id, data_id = model_id), size=1, show.legend = FALSE)  +
  labs(title = "Forecast skill")

## caption doesn't display all that nicely, better to use storyboard
      #caption = 
      #   "Top: Forecast skill measured by CRPS and log skill.\\n
      #   Lower values indicate better predictions. Both score\\n
      #   probablistic forecasts, but log skill penalizes     \\n
      #   observations outside expected range much more heavily.")
```

Overall model_id rankings by forecast skill, as well as skill by date of initial forecast.

**Top plots**: Forecast skill measured by CRPS and log skill. Lower values indicate better predictions. Both score probabilistic forecasts, but log skill penalizes observations outside expected range much more heavily.

**Lower plot**: Many phenology teams submit a new 35-day forecast daily. This plot summarizes the score of each of those 35-day forecasts (over all 35 days and over all sites) as the challenge progresses.

**Tip**: *Mouse over a color to highlight the scores of one team*

```{r girafe, fig.width=8, fig.height=4}
ggob <- board1 / board2 # patchwork stack
girafe(ggobj = ggob,
       width_svg = 8, height_svg = 4,
  options = list(
    opts_hover_inv(css = "opacity:0.20;"),
    opts_hover(css = "stroke-width:2;"),
    opts_zoom(max = 4)
  ))
```

------------------------------------------------------------------------

```{r}
## more possible stuff, not plotted
some_sites <- c("BART", "HARV", "WOOD", "DELA")
greenness_forecasts <-
  combined |>
  filter(reference_datetime==max(reference_datetime, na.rm=TRUE), site_id %in% some_sites) |>
  filter(variable == "gcc_90") |> 
  ggplot() +
  geom_point(aes(time, observation), size = .05) + 
  geom_ribbon_interactive(aes(x = time, ymin = quantile02.5, ymax = quantile97.5,
                              fill = model_id, data_id = model_id, tooltip = model_id),
                          alpha = 0.2, show.legend=FALSE) +
  geom_line_interactive(aes(time, mean, col = model_id, 
                            tooltip = model_id, data_id = model_id), show.legend=FALSE) +
  labs(y = "greenness index (gcc 90)") +
  facet_wrap(~site_id, ncol=2) + 
  theme(axis.text.x = element_text( angle = 90, hjust = 0.5, vjust = 0.5)) +
  ggtitle("30th day ahead or longer prediction") 
```
