---
title: "Ticks"
---

```{r, setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
source("R/plot-utils.R")
score4cast::ignore_sigpipe()

```

```{r}
theme <- "ticks"
cutoff <- as.character(Sys.Date() - 120)
combined <- arrow::open_dataset(glue("cache/parquet/{theme}")) |>
  filter(date >= cutoff) |> collect() 

#combined |> filter(!is.na(observation))

```

```{r}
knitr::opts_chunk$set(eval=FALSE) # no tick data
```

## Most recent forecasts

```{r}
sites <- combined |> distinct(site_id) |> collect() |> pull(site_id)
## with at least n observations to compare!
n_data <- 30
who <- combined |> 
 # filter(!is.na(observation)) |> 
  distinct(reference_datetime) |> 
  arrange(desc(reference_datetime)) |> 
  collect()
ref <- as.character (who[6,1]) # 
```

### *Amblyomma americanum*

::: panel-tabset
```{r}
richness <- combined |> 
  filter(reference_datetime == ref, variable == "amblyomma_americanum")
```

## Sites 1 - 9

```{r}
richness |>  filter(site_id %in% sites[1:9]) |> forecast_plots(ncol=4)
```
:::

## Leaderboard

Average skill scores of each model across all sites.\
Scores are shown by reference date and forecast horizon (in days).\
Scores are averaged across all submissions of the model with a given horizon or a given `reference_datetime` out of submissions made since the cutoff date.

```{r include=TRUE}
cutoff
```

::: panel-tabset
## *Amblyomma americanum*

```{r}
leaderboard_plots(combined, "amblyomma_americanum")

```
:::

## Submission statistics

```{r}
n_models <- combined |> distinct(model_id) |> nrow()
```

Between `r cutoff` and `r Sys.Date()`:

-   `r n_models` submitted to this challenge

```{r}
current_day <- Sys.Date()

combined |> 
  distinct(model_id, reference_datetime) |>
  count(model_id, sort=TRUE) |>
  filter(n > 1) |>
  mutate(model_id = fct_reorder(model_id, n)) |>
  ggplot(aes(model_id, n, fill=model_id)) +
  geom_col(show.legend = FALSE) + 
  coord_flip() + 
  theme_bw() + 
  theme(axis.text=element_text(size=6)) +
  ggtitle(glue("Number of forecast submissions since {cutoff}"))
```
